Hydra Battles and Cie

### 2021-08-03 doc/part-hydras.tex
converting chapter 3 to alectryon (continued)

\begin{Coqsrc}
Inductive CompareSpec (Peq Plt Pgt : Prop) : comparison -> Prop :=
    CompEq : Peq -> CompareSpec Peq Plt Pgt Eq
  | CompLt : Plt -> CompareSpec Peq Plt Pgt Lt
  | CompGt : Pgt -> CompareSpec Peq Plt Pgt Gt
\end{Coqsrc}

\begin{Coqsrc}
Class Comparable {A:Type} 
  (lt: relation A)
  (compare : A -> A -> comparison) :=
  {
  sto :> StrictOrder lt;
  compare_correct: forall a b,
      CompareSpec (a = b) (lt a b) (lt b a) (compare a b);
  }.
\end{Coqsrc}

@@

\begin{Coqsrc}
Class ON {A:Type}(lt: relation A)
      (compare : A -> A -> comparison)  :=
  {
  comp :> Comparable lt compare;
  wf : well_founded lt;
  }.
\end{Coqsrc}

@@

\label{sect:on-lt-notation}
\label{sect:on-le-notation}
\label{sect:measure-ON}
\begin{Coqsrc}
Section Definitions.

  Context  {A:Type}{lt : relation A}
           {compare : A -> A -> comparison}
           {on : ON lt compare}.

  #[using="All"]
   Definition ON_t := A.

  #[using="All"]
   Definition ON_compare := compare.

  #[using="All"]
   Definition ON_lt := lt.

  #[using="All"]
   Definition ON_le:  relation A := leq lt.

  #[using="All"]
   Definition measure_lt {B : Type} (m : B -> A) : relation B :=
    fun x y =>  ON_lt (m x) (m y).

  #[using="All"]
   Lemma wf_measure {B : Type} (m : B -> A) :
    well_founded (measure_lt m).

  #[using="All"]
   Definition ZeroLimitSucc_dec :=
    forall alpha,
      {Least alpha} +
      {Limit alpha} +
      {beta: A | Successor alpha beta}.

  (** The segment called [O alpha] in Schutte's book *)

  #[using="All"]
   Definition bigO 
   (a: A) : Ensemble A := fun x: A => lt x a.

End Definitions.

Infix "o<" := ON_lt : ON_scope.
Infix "o<=" := ON_le : ON_scope.
Infix "o?=" := ON_compare (at level 70) : ON_scope.
\end{Coqsrc}

@@

\begin{Coqsrc}
#[global]
 Instance Omega_comp : Comparable Peano.lt  Nat.compare.
Proof.
  split.
  - apply Nat.lt_strorder.
  - apply Nat.compare_spec.
Qed.

#[global]
 Instance Omega : ON  Peano.lt  Nat.compare.
Proof.
 split.
 - apply Omega_comp.
 - apply Wf_nat.lt_wf.
Qed.

#[local] Open Scope ON_scope.
Compute 6 o?= 9.
\end{Coqsrc}

@@

Definition t := (A + B)%type.
Arguments inl  {A B} _.
Arguments inr  {A B} _.

Definition lt : relation t := le_AsB _ _ ltA ltB.
\end{Coqsrc}

In order to build an instance of \texttt{ON}, we have to define a comparison function.

\begin{Coqsrc}
Definition compare (alpha beta: t) : comparison :=
   match alpha, beta with
     inl _, inr _ => Lt
   | inl a, inl a' => compareA a a'
   | inr b, inr b' => compareB b b'
   | inr _, inl _ => Gt
  end.

#[global] Instance plus_comp : Comparable lt compare.
Proof.
  split.
  - apply lt_strorder.
  - apply compare_correct. 
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma lt_wf : well_founded lt.
Proof. 
  destruct NA, NB; apply wf_disjoint_sum; [apply wf | apply wf0].
Qed.
\end{Coqsrc}

Then, we register an instance of \texttt{ON}:

\begin{Coqsrc}
#[global] Instance ON_plus : ON lt compare.
Proof.
  split.
  - apply plus_comp.
  - apply lt_wf.
Qed.
\end{Coqsrc}

@@

The ordinal $\omega+\omega$ (also known as $\omega\times 2$) may be represented as the concatenation 
of two copies of $\omega$ (Figure~\ref{fig:omega-plus-omega}).

@@

\begin{Coqsrc}
Instance Omega_plus_Omega : ON _ _ := ON_plus Omega Omega.

Definition t := ON_t.
\end{Coqsrc}

\begin{Coqsrc}
Example ex1 : inl 7 o< inr 0.
Proof. constructor. Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Definition fin (i:nat) : t := inl i.
Coercion fin : nat >-> t.

Notation omega := (inr  0:t).
\end{Coqsrc}

\begin{Coqsrc}
Compute  fin 8 o?= omega.
\end{Coqsrc}

\begin{Coqsrc}
Lemma lt_omega alpha : 
     alpha o< omega <-> exists n:nat,  alpha = fin n.
(* ... *)
\end{Coqsrc}

@@

\begin{itemize}
\item The least ordinal, \texttt{(inl 0)}, also known as  \texttt{(fin 0)}.
\item The limit ordinal $\omega$.
\item The successor ordinals, either of the form \texttt{(inl (S $i$))} or \texttt{(inr (S $i$))}
\end{itemize}

@@

\begin{Coqsrc}
Section A_given.
  Variables (A : Type)  (lt: relation A).

  #[local] Infix "<" := lt.
  #[local] Infix "<=" := (leq lt).

 Definition Least {sto : StrictOrder lt} (x : A):=
  forall y,  x <= y.

 Definition Successor {sto : StrictOrder lt} (y x : A):=
  x < y /\ (forall z,  x < z ->  z <  y -> False).

 Definition Limit {sto : StrictOrder lt}  (x:A)  :=
   (exists w:A,  w < x) /\
   (forall y:A, y < x -> exists z:A, y < z /\ z < x).
\end{Coqsrc}

>>>>

\begin{Coqsrc}
Inductive CompareSpec (Peq Plt Pgt : Prop) : comparison -> Prop :=
    CompEq : Peq -> CompareSpec Peq Plt Pgt Eq
  | CompLt : Plt -> CompareSpec Peq Plt Pgt Lt
  | CompGt : Pgt -> CompareSpec Peq Plt Pgt Gt.
\end{Coqsrc}

\emph{From Module~\href{../theories/html/hydras.Prelude/mparable.html\#Hvariant}{Prelude.Comparable}}

\input{movies/snippets/Comparable/ComparableDef}

@@

\input{movies/snippets/ON_Generic/ONDef}

@@

\label{sect:on-lt-notation}
\label{sect:on-le-notation}
\label{sect:measure-ON}

\input{movies/snippets/ON_Generic/ONDefs}

@@

\input{movies/snippets/ON_Omega/OmegaDef}

@@

\input{movies/snippets/ON_plus/Defs}

In order to build an instance of \texttt{Comparable}, we have to define a correct comparison function.

\input{movies/snippets/ON_plus/compareDef}

\input{movies/snippets/ON_plus/compareCorrect}

\input{movies/snippets/ON_plus/plusComp}

@@

\input{movies/snippets/ON_plus/ltWf}

Then, we can build an instance of \texttt{ON}:

\input{movies/snippets/ON_plus/OnPlus}

@@

The ordinal $\omega+\omega$ (also known as $\omega\times 2$) may be represented as the concatenation 
of two copies of $\omega$ (Figure~\ref{fig:omega-plus-omega}).
It is also represented by the two first lines of Figure~\ref{fig:ordinal-sequence}.

@@

\input{movies/snippets/ON_Omega_plus_omega/OmegaPlusOmegaDef}

@@

\input{movies/snippets/ON_Omega_plus_omega/finiteOmega}

\input{movies/snippets/ON_Omega_plus_omega/ltOmega}

@@

\begin{itemize}
\item The least ordinal, \texttt{(inl 0)}, also known as  \texttt{(fin 0)}.
\item The first infinite ordinal $\omega$.
\item The remaining ordinals, either of the form \texttt{(inl (S $i$))} or \texttt{(inr (S $i$))} (in black on Figure~\ref{fig:ordinal-sequence}), called \emph{successor ordinals}.
\end{itemize}

@@

\input{movies/snippets/MoreOrders/Defs}
